---
title: "Traversing the Hydrologic Network"
author: "Vignette Author"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Traversing the Hydrologic Network}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Find a place in the network to start

Lets say we have an observation `lakeobs1` that was taken at some point. We want to see
if there is a lake at that point, and get the information, including ID, of that lake. 

```{r eval=TRUE, echo=TRUE, warning=FALSE, include=FALSE}
library(hydrolinks)
library(sf)
linked_wb = link_to_waterbodies(39.086637, -120.052634, 'lakeobs1', dataset='nhdh')
```

Lo and behold, this point is over top of Lake Tahoe. We can get further information
about this lake from the mapping layer we used to link. National Hydrography Dataset
Highres (NHDH) has some other useful information included. 

```{r eval=TRUE, echo=FALSE}
print(linked_wb)
```

Now, lets generate map of the linked lake (Lake Tahoe). Using hydrolinks, we don't need
to dig through shapefiles. We can just use the ID we got from linking. 

```{r eval=TRUE, echo=TRUE, message=FALSE, warning=FALSE, results='hide'}
lake_poly = get_shape_by_id(linked_wb$permanent_, dataset = 'nhdh', feature_type = 'waterbody')
#Simple quick viz of polygon
plot(st_geometry(lake_poly), main='Lake Tahoe', col='Dodgerblue')

```

The polygon returned is fully functional, so it could be used for 
quantiative analysis of the lake shape or location. We use the `sf` package
throughout for spatial data handling.

## Network functionality

The latest and greatest feature allows for the traversal of the hydrologic network. 
Buildling on what we did above, we can quickly grab very useful information from 
the hydrologic network. 

### Upstream traversal

For example, lets grab and plot all inflows into Lake Tahoe. We will start from
the lake and traverse up the hydrologic network. We will specify a max traversal
distance of 50km to prevent the traversal from exploding (can happen with unlimited
bounds going up the network, e.g., Mississippi). 

```{r eval=TRUE, echo=TRUE, message=FALSE, warning=FALSE, results='hide'}

upstream = traverse_flowlines(50, linked_wb$permanent_, direction = 'in')
upstream_shp = get_shape_by_id(upstream$permanent_, dataset = 'nhdh', feature_type = 'flowline')
plot(st_geometry(upstream_shp), col='palegreen')
plot(st_geometry(lake_poly), main='Lake Tahoe', col='Dodgerblue', add=TRUE)

```

Boom! We have Tahoe and all its input tributaries. 
Again, the shape data returned are the same as from the underlying hydrologic 
network datasets, so they can be used in spatial and other analyses. 

**Note:** Upstream traversal can quickly blow up if you are not careful. 
The river network is very dendritic and so the traversal quickly becomes 
expontential the further up any stream network you go (unless you are near the
headwater streams).

### Downstream traversal

We can also do downstream traversal. We'll start at Lake Tahoe above and see 
where the water goes. One of the unique aspects of this is we have combined
the flow network for both lakes and streams. This means when we do the traversal, we
can examine the stream and lake portion of the flow network independently. Below is an
example using Tahoe, which, I did not earlier know, doesn't drain to the ocean. 

```{r eval=TRUE, echo=TRUE, message=FALSE, warning=FALSE, results='hide'}

downstream = traverse_flowlines(2000, linked_wb$permanent_, direction = 'out')
downstream_shp = get_shape_by_id(downstream$permanent_, dataset = 'nhdh', feature_type = 'flowline')
downstream_lk_shp = get_shape_by_id(downstream$permanent_, dataset = 'nhdh', feature_type = 'waterbody')
plot(st_geometry(downstream_shp), col='palegreen')
plot(st_geometry(downstream_lk_shp), main='Lake Tahoe', col='Dodgerblue', add=TRUE)
plot(st_geometry(lake_poly), main='Lake Tahoe', col='Dodgerblue', add=TRUE)

```


### Another Example

Here is another example of Lake Mendota, Wisconsin. This does an upstream traversal. You can 
see the multiple small impoundments and other ponds/lakes in the upstream network. 

```{r eval=TRUE, echo=TRUE, message=FALSE, warning=FALSE, results='hide'}
id = link_to_waterbodies(43.112449, -89.429409, 'mendota')

hmm = traverse_flowlines(100, id$permanent_, "in")

fls = get_shape_by_id(hmm$permanent_, feature_type = 'flowline', dataset='nhdh')
wbs = get_shape_by_id(hmm$permanent_, feature_type = 'waterbody', dataset='nhdh')

wb   = get_shape_by_id(id$permanent_, feature_type = 'waterbody')

plot(st_geometry(fls), col='green')
plot(st_geometry(wb), col='orange', add=TRUE)
plot(st_geometry(fls), add=TRUE, col='green')
plot(st_geometry(wbs), add=TRUE, col='blue')

111
```
