---
title: "WQP Lake Linking Example"
author: "Luke Winslow"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{WQP Lake Linking Example}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache=TRUE)
```

## Linking Water Quality Portal 

This is an example of using `hydrolinks` to link a large, unlinked aquatic dataset to the U.S. Geological 
Survey's National Hydrography Network. The unlinked aquatic dataset is from the Water Quality Portal (WQP). 
In this example, I will link a lakes, but you could also link streams. 

The WQP data is a perfect use-case here. The data are all described by sites that have Latitude/Longitude geopoint
data, but no other concrete way to link the data to the hydrologic network. So in this example, we will link
all lake observation sites in the WQP to the National Hydrography Dataset high-res (NHDH). Any sites that don't link
using the simple point-in-polygon routine 

First, load packages and grab lake data. I'm limiting the data to be from the state of Wisconsin initially. 
This just speeds up generation of the Vignette. The entire US dataset takes a bit longer than I'd like to wait
to link and visualize. 

```{r packages_sites, eval=TRUE, echo=TRUE, warning=FALSE, include=TRUE, message=FALSE}
library(dataRetrieval)
library(hydrolinks)
library(USAboundaries)
library(ggplot2)
library(gridExtra)
library(sf)

lsites = whatWQPsites(siteType ="Lake, Reservoir, Impoundment", statecode='US:55')
```

At the point of this Vignette generation, there were `r length(unique(lsites$MATCH_ID))` unique sites 
returned by this query. 

Now, lets link all these sites to the NHDH. After the initial point-in-polygon linking, I will take
the remainder and use the centroid matching. 

```{r link, eval=TRUE, echo=TRUE, warning=FALSE, include=TRUE, message=FALSE}

linked_lakes = link_to_waterbodies(lsites$LatitudeMeasure, lsites$LongitudeMeasure, lsites$MonitoringLocationIdentifier, dataset = 'nhdh', buffer = 0)

unlinked = subset(lsites, !(MonitoringLocationIdentifier %in% linked_lakes$MATCH_ID))
#now, try linking just the previously unlinked sites
linked_lcent = link_waterbody_centroids(unlinked$LatitudeMeasure, unlinked$LongitudeMeasure, unlinked$MonitoringLocationIdentifier, dataset = 'nhdh', max_dist = 25)
```


Here are the overarching number of sites linked/centroid-linked/not-linked. 

```{r build_table, echo=FALSE, message=FALSE, warning=FALSE}
library(knitr)
library(kableExtra)
tbl = data.frame(Link=c('Point-in-polygon', 'Centroid', 'Unlinked', 'Total'), 
                 Count=c(length(unique(linked_lakes$MATCH_ID)), length(unique(linked_lcent$MATCH_ID)), nrow(subset(lsites, !(MonitoringLocationIdentifier %in% linked_lakes$MATCH_ID) & !(MonitoringLocationIdentifier %in% linked_lcent$MATCH_ID))), length(unique(lsites$MonitoringLocationIdentifier))))

kable(tbl, 'html') %>%   kable_styling()
```

Lastly, lets visualize these sites. Little `grid.arrange` to mash them all together. 

```{r mapfigs, echo=TRUE, warning=FALSE, message=FALSE, fig.width=8, fig.height=8}
wi = USAboundaries::us_boundaries()[USAboundaries::us_boundaries()$geoid %in% c('55'),]
all_points = st_as_sf(lsites, coords = c("LongitudeMeasure", "LatitudeMeasure"), crs = '+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs')


gall = ggplot(all_points) + geom_sf(data=wi, fill='grey') +
  geom_sf(size=0.5, color='darkorange2') + ylim(42, 47) + xlim(-93, -86.5) +
  ggtitle('a) all WQP points')

glink = ggplot(subset(all_points, MonitoringLocationIdentifier %in% linked_lakes$MATCH_ID)) +
  geom_sf(data=wi, fill='grey') +
  geom_sf(size=0.5, color='dodgerblue') + ylim(42, 47) + xlim(-93, -86.5) +
  ggtitle('b) all linked point-in-poly')

gcent = ggplot(subset(all_points, MonitoringLocationIdentifier %in% linked_lcent$MATCH_ID)) +
  geom_sf(data=wi, fill='grey') +
  geom_sf(size=0.5, color='darkorchid3') + ylim(42, 47) + xlim(-93, -86.5) +
  ggtitle('c) all linked centroid')


grid.arrange(gall, glink, gcent, nrow = 2)
```


